A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2018-09-29 paper/submission/figures/opt-framework_code.tex
adding figures to submission folder

	\path (pilot.north)+(0.7,1.5) node (opal) [draw, text centered,
	text width=6em, fill=orange, fill opacity=0.1, minimum height=2em,
	rounded corners, very thick, draw=orange, text opacity=1.0]
	{\textsc{Simulation}};

>>>>

	\path (pilot.north)+(0.7,1.5) node (opal) [draw, text centered,
	text width=8em, fill=orange, fill opacity=0.1, minimum height=2em,
	rounded corners, very thick, draw=orange, text opacity=1.0]
	{\textsc{Beam Dynamics}};

### 2018-09-29 paper/submission/prab.tex
adding figures to submission folder

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research but as well as practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Today, tuning machine parameters, e.g. bunch charge, emission time and
  various parameters of beamline elements, is most commonly done manually by
  running simulation codes to scan the parameter space.
This approach is tedious, time consuming and can be error prone.
In order to be able to reliably identify optimal configurations of
  accelerators we propose to solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning
  parameters.
Observe that multiple and conflicting optimality criteria call for a
  multi-objective approach.

>>>>

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research but as well as practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Today, tuning machine parameters, e.g. bunch charge, emission time and
  various parameters of beam line elements, is most commonly done manually by
  running simulation codes to scan the parameter space.
This approach is tedious, time consuming and can be error prone.
In order to be able to reliably identify optimal configurations of
  accelerators we propose to solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning
  parameters.
Observe that multiple and conflicting optimality criteria call for a
  multi-objective approach.

### 2018-09-29 paper/submission/theor-experiment.tex
adding figures to submission folder

\def \gunheight {25}
\def \quadheight {35}
\def \quadleft {127}
\def \kickerleft {160}
\def \leftarrow{10}
\def \bottomofarrows {29}
\def \topofarrows {34}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig1) at (0,0)
	{\includegraphics[width=\textwidth]{Report/awa-drawing}};
	\node[fill=white, inner sep=2pt] (txt2) at (0,\gunheight) {\large Gun};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\gunheight) {\large Accelerating Cavities};
	\node[fill=white, inner sep=2pt] (txt2) at (\quadleft,\quadheight) {\large Quadrupoles};
	\draw [blue, ->, line width=1] (\quadleft+\leftarrow,\topofarrows) -- (\quadleft+\leftarrow, \bottomofarrows);
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,20) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft+8,20) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,\quadheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kickerleft+\leftarrow-3,\topofarrows) -- (\kickerleft+\leftarrow-3, \bottomofarrows);	
	\draw [black, ->, line width=1] (0,-2) -- (100, -2);
	\node[fill=white, inner sep=2pt] (txt2) at (50,0) {\large Beam Direction};
	\end{tikzpicture}
	\caption{Side view of the high charge linac at the AWA. 
		All hardware in this drawing is currently installed. 
	Note locations $s_1$ and $s_2$, before and after the kicker.}
	\label{awa-linac}
	\vspace{-7em}
\end{figure*}

\def \septheight {65}
\def \bottomarrow {57}
\def \kleft {13}
\def \sleft {41.5}
\def \sheight {47}
\def \beam {35}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{Report/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (46,45) {\large $s_3$};
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-12em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed. 
	Note $s_3$, the entrance to the fifth quadrupole on the beam line.
	This is the optimization location.}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

@@

\begin{figure}%[h]
	\centering
	\includegraphics[width=\linewidth]{Report/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

\iffalse
\begin{table}%[h!]
	\begin{center}
		\caption{Low and mid fidelity simulation parameters.}
		\label{fidelity}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}} C c D }
			\hline\noalign{\smallskip}
			& Number of particles & dT (seconds) & Time to simulation (minutes)\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			low fidelity  			&  10,000   & $5 \times10^{-11}$  &  2.6 \\
			adjusted low fidelity  	&  10,000   & $2 \times 10^{-12}$, $1 \times 10^{-11}$ & 1.6 \\
			mid fidelity 			&  100,000  & $1 \times 10^{-12}$ &  18\\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}
\fi

@@

\begin{figure}
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		(a) Pareto fronts for ex-1 through ex-4 at $s_1$.
		%\caption{(a) Pareto fronts at $s_1$.}
		%\label{expareto1}
		\vspace{2em}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto2}
		(b) Pareto fronts for ex-1 through ex-4 at $s_2$.
		%\caption{(b) Pareto fronts at $s_2$.}
		%\label{expareto2}
	\end{minipage}
	\caption{Comparison of Pareto fronts for initial optimization experiments, ex-1 through ex-4.}
	\label{expareto}
\end{figure}

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$ [m]. 
\begin{figure}
	\begin{center}		
		\begin{tikzpicture}[scale=1]
		\node (fig1) at (0,0)
		{\includegraphics[width=0.5\textwidth]{../pareto_stat_plots/xy_vs_pxy_pareto_front_quads_before_Q5}};
		\node[star, star points=5, star point ratio=0.5, fill=yellow, draw](key) at (3.35,-1.7) {};
		\end{tikzpicture}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes ($\sigma_{x,y}$) and transverse momentum ($\sigma_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

\iffalse
\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{../pareto_stat_plots/dE_vs_zrms_pareto_front_quads_before_Q5_zoomout}%
	\llap{\raisebox{1.9cm}{%  move next graphics to top right corner
			\includegraphics[height=4.5cm]{../pareto_stat_plots/dE_vs_zrms_pareto_front_quads_before_Q5}%
	}}
	\caption{Pareto front comparing energy spread (dE) and bunch length ($\sigma_z$). \nrnote{add tikz arrows to point to left side of curve?}}
	\label{fig:pareto2}
\end{figure}  
\fi

@@

\begin{figure}
	\includegraphics[width=0.48\textwidth]{../pareto_stat_plots/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for large scale TBA optimization runs.}
		\label{tab:designopt}   
		\begin{tabular}{l|c|c}
			\noalign{\smallskip}\hline\noalign{\smallskip}
			%\rowcolor{blue!30} 
			\textbf{Design Variable} & \textbf{Unit}	&  \textbf{Value}  \\ 
			\noalign{\smallskip}\hline\noalign{\smallskip}
			{Buck Focusing Solenoid} & amps	& 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
		\end{tabular}
	\end{center}
\end{table}

>>>>

\def \gunheight {25}
\def \quadheight {35}
\def \quadleft {127}
\def \kickerleft {160}
\def \leftarrow{10}
\def \bottomofarrows {29}
\def \topofarrows {34}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig1) at (0,0)
	{\includegraphics[width=\textwidth]{figures/awa-drawing}};
	\node[fill=white, inner sep=2pt] (txt2) at (0,\gunheight) {\large Gun};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\gunheight) {\large Accelerating Cavities};
	\node[fill=white, inner sep=2pt] (txt2) at (\quadleft,\quadheight) {\large Quadrupoles};
	\draw [blue, ->, line width=1] (\quadleft+\leftarrow,\topofarrows) -- (\quadleft+\leftarrow, \bottomofarrows);
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,20) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft+8,20) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,\quadheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kickerleft+\leftarrow-3,\topofarrows) -- (\kickerleft+\leftarrow-3, \bottomofarrows);	
	\draw [black, ->, line width=1] (0,-2) -- (100, -2);
	\node[fill=white, inner sep=2pt] (txt2) at (50,0) {\large Beam Direction};
	\end{tikzpicture}
	\caption{Side view of the high charge linac at the AWA. 
		All hardware in this drawing is currently installed. 
	Note locations $s_1$ and $s_2$, before and after the kicker.}
	\label{awa-linac}
	\vspace{-7em}
\end{figure*}

\def \septheight {65}
\def \bottomarrow {57}
\def \kleft {13}
\def \sleft {41.5}
\def \sheight {47}
\def \beam {35}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{figures/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (46,45) {\large $s_3$};
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-12em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed. 
	Note $s_3$, the entrance to the fifth quadrupole on the beam line.
	This is the optimization location.}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

@@

\begin{figure}%[h]
	\centering
	\includegraphics[width=\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/ex-pareto1}
		(a) Pareto fronts for ex-1 through ex-4 at $s_1$.
		%\caption{(a) Pareto fronts at $s_1$.}
		%\label{expareto1}
		\vspace{2em}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/ex-pareto2}
		(b) Pareto fronts for ex-1 through ex-4 at $s_2$.
		%\caption{(b) Pareto fronts at $s_2$.}
		%\label{expareto2}
	\end{minipage}
	\caption{Comparison of Pareto fronts for initial optimization experiments, ex-1 through ex-4.}
	\label{expareto}
\end{figure}

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$ [m]. 
\begin{figure}
	\begin{center}		
		\begin{tikzpicture}[scale=1]
		\node (fig1) at (0,0)
		{\includegraphics[width=0.5\textwidth]{figures/xy_vs_pxy_pareto_front_quads_before_Q5}};
		\node[star, star points=5, star point ratio=0.5, fill=yellow, draw](key) at (3.35,-1.7) {};
		\end{tikzpicture}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes ($\sigma_{x,y}$) and transverse momentum ($\sigma_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.48\textwidth]{figures/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for large scale TBA optimization runs.}
		\label{tab:designopt}   
		\begin{tabular}{l|c|c}
			\noalign{\smallskip}\hline\noalign{\smallskip}
			%\rowcolor{blue!30} 
			\textbf{Design Variable} & \textbf{Unit}	&  \textbf{Value}  \\ 
			\noalign{\smallskip}\hline\noalign{\smallskip}
			{Buck Focusing Solenoid} & amps	& 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
		\end{tabular}
	\end{center}
\end{table}
